R Packages to Aid in Handling Web Access Logs

### 2015-12-20 paper/paper.Rmd
a bit more

Base R does not have native code to parse URLs, but the **httr** package \citep{httr} contains a function, `parse_url`, designed to do just that and built on base R's capabilities. This function is neither vectorised nor based on compiled code, and produces a list, making looping over a set of URLs to parse each one a somewhat thorny experience. This is totally understandable given the intent behind that function, which is to decompose individual URLs within the context of making HTTP requests, rather than to analyse request logs *en masse*.

`urltools` contains `url_parse` - which does the same thing as the equivalent `httr` functionality, but in a vectorised way, relying on compiled code, and producing a data.frame. Within the context of parsing and processing access logs, this is far more useful, because it works efficiently over large sets: httr's functionality, not having designed to be, does not. Indeed, benchmarking showed that httr's version is (for large vectors, at least) approximately 570 times slower than `url_parse`:

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.4]{decoding_benchmarks}
    \caption{Results of microbenchmark run: url\char`_parse versus httr equivalent code}
\end{figure}

>>>>

Base R does not have native code to parse URLs, but the **httr** package \citep{httr} contains a function, `parse_url`, designed to do just thatm built on R's regular expressions. This function is not vectorised, does not make use of compiled code internally, and produces a list rather than data.frame, making looping over a set of URLs to parse each one a time-consuming and awkward experience. This is understandable given the intent behind that function, which is to decompose individual URLs within the context of making HTTP requests, rather than to analyse URLs *en masse*.

`urltools` contains `url_parse` - which does the same thing as the equivalent `httr` functionality, but in a vectorised way, relying on compiled code, and producing a data.frame. Within the context of parsing and processing access logs, this is far more useful, because it works efficiently over large sets: httr's functionality, not having designed to be, does not. Indeed, benchmarking showed that `url_parse` is approximately 570 times faster than httr's equivalent function:

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.4]{parsing_benchmarks}
    \caption{Results of microbenchmark run: url\char`_parse versus httr's equivalent code}
\end{figure}

\newpage

A vector of URLs passed into `url_parse` produces a data.frame, with one column for each of the IETF-supported components, and empty strings representing components that could not be found. Additionally, influenced by the style of the *lubridate* package \citep{lubridate}, `urltools` contains functions to get or *set* individual components:

```{r, eval=TRUE}
# Load urltools and construct a URL
library(urltools)
url <- "http://www.google.com/"

# Get the scheme
scheme(url)
```

```{r, eval=TRUE}
# Set the scheme and observe the modification to the resulting URL
scheme(url) <- "https"
url
```
## IP manipulation and geolocation

### 2015-12-20 paper/paper.tex
a bit more

Base R does not have native code to parse URLs, but the \textbf{httr}
package \citep{httr} contains a function, \texttt{parse\_url}, designed
to do just that and built on base R's capabilities. This function is
neither vectorised nor based on compiled code, and produces a list,
making looping over a set of URLs to parse each one a somewhat thorny
experience. This is totally understandable given the intent behind that
function, which is to decompose individual URLs within the context of
making HTTP requests, rather than to analyse request logs \emph{en
masse}.

\texttt{urltools} contains \texttt{url\_parse} - which does the same
thing as the equivalent \texttt{httr} functionality, but in a vectorised
way, relying on compiled code, and producing a data.frame. Within the
context of parsing and processing access logs, this is far more useful,
because it works efficiently over large sets: httr's functionality, not
having designed to be, does not. Indeed, benchmarking showed that httr's
version is (for large vectors, at least) approximately 570 times slower
than \texttt{url\_parse}:

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.4]{decoding_benchmarks}
    \caption{Results of microbenchmark run: url\char`_parse versus httr equivalent code}
\end{figure}

>>>>

Base R does not have native code to parse URLs, but the \textbf{httr}
package \citep{httr} contains a function, \texttt{parse\_url}, designed
to do just thatm built on R's regular expressions. This function is not
vectorised, does not make use of compiled code internally, and produces
a list rather than data.frame, making looping over a set of URLs to
parse each one a time-consuming and awkward experience. This is
understandable given the intent behind that function, which is to
decompose individual URLs within the context of making HTTP requests,
rather than to analyse URLs \emph{en masse}.

\texttt{urltools} contains \texttt{url\_parse} - which does the same
thing as the equivalent \texttt{httr} functionality, but in a vectorised
way, relying on compiled code, and producing a data.frame. Within the
context of parsing and processing access logs, this is far more useful,
because it works efficiently over large sets: httr's functionality, not
having designed to be, does not. Indeed, benchmarking showed that
\texttt{url\_parse} is approximately 570 times faster than httr's
equivalent function:

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.4]{parsing_benchmarks}
    \caption{Results of microbenchmark run: url\char`_parse versus httr's equivalent code}
\end{figure}

\newpage

A vector of URLs passed into \texttt{url\_parse} produces a data.frame,
with one column for each of the IETF-supported components, and empty
strings representing components that could not be found. Additionally,
influenced by the style of the \emph{lubridate} package
\citep{lubridate}, \texttt{urltools} contains functions to get or set
\emph{individual} components:

\begin{Schunk}
\begin{Sinput}
library(urltools)
url <- "http://www.google.com/"
scheme(url)
\end{Sinput}
\begin{Soutput}
#> [1] "http"
\end{Soutput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
scheme(url) <- "https"
url
\end{Sinput}
\begin{Soutput}
#> [1] "https://www.google.com/"
\end{Soutput}
\end{Schunk}

\subsection{IP manipulation and
geolocation}\label{ip-manipulation-and-geolocation}
